---
title: "Practical Machine Learning - Course Project"
author: "Francisco Estrada"
date: "9 de septiembre de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Download the data

```{r, include=FALSE}
library(knitr)
library(caret)
library(SparseM)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
library(e1071)
set.seed(12345)
```

```{r}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))
```

#Create a partition

```{r}
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
dim(TestSet)
```

#Cleaning Data

```{r}
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
dim(TestSet)

AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
dim(TestSet)

TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```

#Correlation Analysis
```{r}
corMatrix <- cor(TrainSet[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```

# 1) Method: model fit

```{r}
RF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=RF)
modRandForest$finalModel

# prediction
predictRandForest <- predict(modRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

#2) Method: Decision Trees

```{r}
modDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modDecTree)

# prediction
predictDecTree <- predict(modDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree
```

#3) Method: Generalized Boosted Model

```{r}
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel

# prediction
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM
```


#Applying the Selected Model 

The accuracy of the 3 models is
Random Forest : 0.9963
Decision Tree : 0.7368
Generalized Boosted Model : 0.9839

the best model us Random Forest and is used for testing data

```{r}
TEST <- predict(modRandForest, newdata=testing)
TEST
```

